%!TEX root = doc.tex
\section{Conclusion}
\label{sec:conclusion}

The development of MAS may require the use of tools that are not readily available in the agent-based framework chosen for the task. The API architecture we propose in this paper emerges as a need to solve a concrete instance of this problem. \apiname{} is our solution to integrating FIPA specifications in MABS frameworks, more specifically Repast, as well as allowing the creation of JADE-like simulations.

The goal of our API is not only to enrich Repast and other simulation frameworks, but also to allow proficient JADE developers in quickly getting started in developing simulations using familiar JADE-based features. \apiname{} also opens the door to MAS to MABS conversion.

The short experiment we described in \ref{sec:verification} showed that our API is capable of handling simulations with a large number of agents, which is common in many simulation scenarios. Aside from the inherent advantages of using open standards in agent systems, \apiname{}'s JADE-based architecture brings Repast closer to JADE. The immediate result of this architecture is that converting models built with Repast using \apiname{} into JADE MAS is very straightforward.

Multiple candidates exist that could have been the object of this paper. JADE and Repast were chosen for their large and active community of researchers, as well as for being open source. One important goal of our project is to eventually release our tools (both \apiname{} and the code conversion tool) for use by the cientific community.
